

## Initial Loading 

With Below code, we load libraries needed for the program, set the directory and read the excel file. We also add more custoem columns to the data frame used in next steps 

```{r, echo=TRUE}
library(knitr)
library(ggplot2)
library(xtable)
library(plyr)
library(lubridate)
setwd ( "c:/temp")
data <- read.csv ( "activity.csv", sep = "," , header= TRUE, stringsAsFactors = FALSE, na.strings = "NA")

data$activityDate <- as.Date(data$date, "%Y-%m-%d")
data$day <- wday( data$activityDate)
data$datecategory <- ifelse ( data$day == 1 | data$day == 7 , "Weekend", "WeekDay")
nrow(data)
summary(data)
head(data)
```

## Mean total number of steps taken per day

```{r, echo=TRUE,fig.height=10, fig.width=10}
dailySummary <- ddply ( data, .(activityDate), summarise, total.step = sum(steps, na.rm=TRUE) )


summary( dailySummary$total.step)

hist ( dailySummary$total.step , breaks = 15, col="green", border = "red", 
       freq=TRUE, plot = TRUE , xlab = "Number of Steps Per Day", ylab= "Frequency" , 
       main= "Frequency of Number of Steps taken per Day")

```

```{r, echo=TRUE, results = "hide" }
m <- median(dailySummary$total.step, na.rm = TRUE)


i <- mean(dailySummary$total.step, na.rm = TRUE)
```

The average of total number of steps taken per day is `r i`, while the median was `r m`

## average daily activity pattern?

```{r , echo=TRUE,fig.height=10, fig.width=10}

intervalSummary <- ddply ( data, .(interval), summarise, avg.step = mean(steps, na.rm=TRUE) , 
                                                         max.step = max(steps, na.rm= TRUE))


plot ( intervalSummary$interval, intervalSummary$avg.step , 
       type = "l" , xlab="interval", 
       ylab="Average of Steps", 
       main= "Average of Steps on 5 minutes interval taken across all days ")
```

As Plot shows, the most of actvities are around 9 to 10 morning. 

## What is effect of Missing data in dataset

```{r missingdata, echo=TRUE, fig.height=10, fig.width=10 }

q <- colSums ( is.na (data))

q

head(data)
```

We populate missing steps with average of steps taken on that interval across all days. 

```{r fillmissingdata, echo=TRUE, fig.height=10, fig.width=10 }

data1 <-merge ( data, intervalSummary, by.x="interval" , by.y="interval" )



data1$steps[is.na(data1$steps)] <-  data1$avg.step


dailySummary1 <- ddply ( data1, .(activityDate), summarise, total.step = sum(steps, na.rm=TRUE) )



hist ( dailySummary$total.step , breaks = 15, col="green", border = "red", 
       freq=TRUE, plot = TRUE , xlab = "Number of Steps Per Day", ylab= "Frequency" , 
       main= "Frequency of Number of Steps taken per Day after missing steps are filled with Avergae steps taken on that interval across all days")

```


```{r, echo=TRUE, results = "hide" }
m1 <- median(dailySummary1$total.step, na.rm = TRUE)

m1

i1 <- mean(dailySummary1$total.step, na.rm = TRUE)

i1
```

####
After filling missing steps with Average steps for that interval across all daya, The average of total number of steps taken per day is `r i1`,
and median was `r m1`

Compared to original dataset, the median hasn't changed, that means, the missing steps are not for frequent intervals, also the mean of dataset hasn't been changed a lot, that means the data that was missing are for intervals that usually there were not much of acitvities.



## Differences in activity patterns between weekdays and weekends
As graphs in bleow represent, during week Days, the average number of steps during morning until 12PM is much higher than weekends.


```{r , echo=TRUE, fig.height=10, fig.width=10 }

qplot( interval, steps, data=data1 , type = "l", facets =  datecategory~., geom= "line")

```
